### 2019-12-12 docs/thesis-project/article.tex
Updates

\subsection
{Frequent approach}There has been much research on the topic of P2P overlays. There are two significant categories of overlays, Structured and Unstructured. In the former, the overlay has rigid rules regarding peer placement; they favor lookup speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay. Some popular P2P Structured overlays include Chord \cite{chord}, Tapestry \cite{tapestry}, Kademlia \cite{kademlia}, and P-Grid \cite{pgrid}. In the latter, peer placement is random, which provides the overlay network with a higher degree of resilience and robustness in the advent of failures. Unstructured overlays disseminate information in a gossip-like fashion; thus, lookup operations are slower, resource management is less efficient, but overlay maintenance is often straightforward. Popular  Unstructured P2P overlays include FreeNet \cite{freenet}, Gnutella \cite{gnutella-rfc} and BitTorrent \cite{bittorrent}. Research on multi-layer overlays has not yielded satisfactory results, and implementation is difficult \cite{sotart}, but research on bio-inspired overlays did have promising results, some of these include Self-Chord \cite{selfchord}. Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify \cite{spotify} and Wuala \cite{wuala}. From all the mentioned research, perhaps the one that borrows most similarity with our proposal is CLOUDCAST; however, neither it nor any of the remaining literature, try to create a structured overlay composed of completely autonomous peers, who work together to keep the file alive without enforced persistence based on PSG.

\subsection{Proposal Goals}\label{subsec:intro}
\begin{itemize}
    \item Implement a Tracker server with light-weight responsibilities:
        \SubItem{Act as peer discovery service for clients;}
        \SubItem{Act as an indexer for clusters of peers (Hives);}
        \SubItem{Track peer reputations;}
    \item Implement an erasure-code algorithm to ensure file persists in the Hive;
    \item Implement a trust system where:
        \SubItem Each peer gives feedback to the Tracker about his neighbors;
        \SubItem Tracker synthesizes the feedback and presents it to the client;
    \item Calculate the ideal steady-state distribution of file blocks, for a Hive, when:
        \SubItem{No trust history on a new peer is available;}
        \SubItem{Peer availability, reliability and other trust factors are known;}
    \item Hives converge to ideal block distribution in bounded-times using PSG;
    \item Efficiently update the PSG's Markov Chain as it changes over time;
    \item Balance the in-degree and out-degree of each peer such that:
        \SubItem{Message exchange overhead is minimal;}
        \SubItem{Any changes to the probability of file loss are negligible;}
        \SubItem{Hive availability and reliability are not compromised;}
        \SubItem{Hive is resilient to massive node failures and resistant to churn;}
    \item Extra miles may include:
        \SubItem{Peers can reconstruct Hives without Tracker intervention;}
        \SubItem{Peers trust their neighbors independently of Tracker;}
        \SubItem{Hives can leverage dedicated Cloud-Storage if they become unreliable;}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
% END INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Related Work}\label{sec:relatedwork}
This section as the following structure: In the first subsection, we will study two distributed file systems used in cloud architectures, in the following two subsections we look at four P2P overlays. We will then mention one cloud-assisted P2P protocol that we consider relevant for being designed to the same use-case that led us to make this thesis proposal. Then we will take a look at two well-known research regarding trust in P2P networks, and finally, we will study erasure code algorithms for file-block replication.
\subsection{Distributed File Systems}
\textbf{Google File System} (GFS) is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers. In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures. Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception \cite{gfs}.  This assumption is typical in modern distributed systems of any kind. GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management. One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks \cite{gfs}. In the GFS architecture cluster consists of a single master and multiple chunk-servers. The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them. Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations. Neither the client nor the chunk-servers cache file data because most of google's client applications stream through huge files. A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs. For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master.

@@

\textbf{Ceph} is an open-source, storage system that, like GFS, runs over commodity hardware. Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure. The base architectural component of Ceph is the RADOS object store, and all content a client reads and writes to the system are represented as objects, irrespective of the objects' original data type. RADOS is for all purposes a Ceph cluster, and its storage space originates from the disks of multiple machines. Ceph architecture is slightly more complicated than that of GFS, so we will summarize the components that are of interest to us. Inside a cluster, we find three types of machines, storage devices, abstracted with a layer called Object Software Daemon (OSD), Monitors, and Metadata Servers. A small count of Monitors maintain cluster membership and state using PAXOS \cite{paxos} to make consensual decisions. Metadata Servers only required sharing the distributed file system, are used for POSIX semantics, e.g., directory hierarchy and file metadata such as owner and last accessed timestamps. Monitors can detect if metadata servers are down and launch new ones as required. Finally, OSDs store clients' data, and there are anywhere between dozens to thousands of these in a Ceph cluster.

@@

\subsection{Structured Overlays}
\textbf{Kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology \cite{kademlia}. Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing. Inserts and lookups on the overlay are unidirectional, according to the distance between two identifiers, calculated using XOR. The unidirectionality allows caching to be done along lookup paths, alleviating hot spots. Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to a least recently seen policy. As a peer receives messages if the sender is a new acquaintance, that sender is placed on the proper k-bucket if there is enough space in it; otherwise, it will replace the least seen peer only if that peer fails to respond to a ping. By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}. This policy also prevents some denial of service attacks, where an enemy floods the network with new peers. The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers in the network to learn useful routing information from discovery and lookup messages. Because of the existence of k-buckets, routing can also be done around failed nodes, providing some resistance to churn and partitioning.

@@

\textbf{Self-Chord} is a bio-inspired, structured overlay, specifically designed to be used in cloud computing or grid computing architectures. Self-chord foundations are that of the standard Chord, both in terms of identifiers and the way peers organize themselves on the overlay, i.e., in ring-like shapes. However, it has a few benefits, which are the result of having mobile agents, called ants, independently roaming the overlay network, but collaboratively re-arranging resource identifiers so that similar data items are clustered around neighboring peers with relatively proportional workloads, meaning that data item placement is not based solely on the numerical-space shared between them and possible placement peers \cite{sotart, selfchord}. Hence advantages of Self-Chord over the standard implementation include lack of need to assign data to well-specified peers since keys are independently defined, leading to better behavior under high churn conditions. Less overlay maintenance overhead since when a peer joins or leaves the network, ants will autonomously and eventually, distribute data items appropriately; Lastly, since load-balancing is continuously being carried out by ants, the scalability and robustness of the system increases since individual peer failures are less likely to lead to massive file loss. According to the authors, the behavior of the ants ensures that even under highly dynamic, unfavorable conditions, rearrangement, and discovery of data items takes only logarithmic time.

@@

\subsection{Unstructured Overlays}
\textbf{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}. Gnutella is the first-ever decentralized, unstructured P2P protocol. Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation. Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive. Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations. The results show that Gia is three to five orders of magnitude better than that of Gnutella. Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent. Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.

@@

\textbf{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols. Like in Self-Chord\cite{selfchord}, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded. The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller, thus achieving the sought goals. BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches,  as an alternative to semantic clustering such as the ones used in Antares\cite{antares} and Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed. The overlay management can be viewed in three major components.

@@

\textbf{CYCLON} \cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays. CYCLON is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn and massive node failures, in a fully decentralized manner. One fundamental issue of unstructured networks is to avoid partitioning. One naive solution would be for each peer to keep a view of all the peers in the network and create connections has needed, but that solution is not scalable.

@@

We conclude by reinforcing the idea that CYCLON is a scalable and robust solution for peer-sampling, providing good self-healing and self-organization capabilities without complex algorithms or apparent downsides. In a way, it relates to BlatAnt since the proposal passively bounds the diameter of the network.

\subsection{Cloud-Assisted P2P Protocol - CLOUDCAST}
\subsection{File Block Replication}
\subsection{Trust in Peer-to-Peer Networks}
\subsection{Probabilistic Swarm Guidance}
%%%%%%%%%%%%%%%%%%%%
% END RELATED WORK %
%%%%%%%%%%%%%%%%%%%%

>>>>

\subsection
{Frequent approach}There has been much research on the topic of P2P overlays. There are two significant categories of overlays, Structured and Unstructured. In the former, the overlay has rigid rules regarding peer placement; they favor lookup speed but have higher maintenance overhead under churn, i.e., the process of peers leaving or joining the overlay. Some popular P2P Structured overlays include Chord \cite{chord}, Tapestry \cite{tapestry}, Kademlia \cite{kademlia}, and P-Grid \cite{pgrid}. In the latter, peer placement is random, which provides the overlay network with a higher degree of resilience and robustness in the advent of failures. Unstructured overlays disseminate information in a gossip-like fashion; thus, lookup operations are slower, resource management is less efficient, but overlay maintenance is often straightforward. Popular  Unstructured P2P overlays include FreeNet \cite{freenet}, Gnutella \cite{gnutella-rfc} and BitTorrent \cite{bittorrent}. Research on multi-layer overlays has not yielded satisfactory results, and implementation is difficult \cite{sotart}, but research on bio-inspired overlays did have promising results, some of these include Self-Chord \cite{selfchord}. Finally, there have also been some attempts of transparently integrating Cloud Systems into P2P networks, e.g., CLOUDCAST \cite{cloudcast}, and vice-versa, e.g., Spotify \cite{spotify} and Wuala \cite{wuala}. From all the mentioned research, perhaps the one that borrows most similarity with our proposal is CLOUDCAST; however, neither it nor any of the remaining literature we present here, try to create a structured overlay composed of completely autonomous peers, who work together to keep the file at all points in time using PSG.

\subsection{Proposal Goals}\label{subsec:intro}
This work seeks to address the problem of file survivability in P2P networks; our primary goal is to study, design, and evaluate a P2P overlay that uses probabilistic swarm guidance as a way to maintain a file available to service clients at all times. Steps will include:

\begin{itemize}
    \item Implement a Tracker server with light-weight responsibilities:
        \SubItem{Act as peer discovery service for clients;}
        \SubItem{Act as an indexer for clusters of peers (Hives);}
        \SubItem{Track peer reputations;}
    \item Implement an erasure-code algorithm used by Hives to replicate clients' files;
    \item Calculate the ideal steady-state distribution of file blocks, for a Hive, when:
        \SubItem{No history information on a peer is available;}
        \SubItem{Peer availability, reliability and other trust factors are known;}
    \item Hives converge to ideal block distribution in bounded-times using PSG;
    \item Efficiently update routing as Hives' topologies change over time;
    \item Balance the degree of each peer such that:
        \SubItem{Message exchange overhead is minimal;}
        \SubItem{Any changes to the probability of file loss are negligible;}
        \SubItem{Hive is resilient to massive node failures and resistant to churn;}
    \item Evaluating the system using PeerSim \cite{peersim} or equivalent platform;
    \item Extra miles may include:
        \SubItem{Implement a trust system managed by the Tracker}
        \SubItem{Peers can reconstruct Hives without Tracker intervention;}
        \SubItem{Peers trust their neighbors independently of Tracker;}
        \SubItem{Hives can leverage dedicated Cloud-Storage if they become unreliable;}
\end{itemize}
%%%%%%%%%%%%%%%%%%%%
% END INTRODUCTION %
%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%
% BEGIN RELATED WORK %
%%%%%%%%%%%%%%%%%%%%%%
\newpage\section{Distributed File Systems}\label{sec:relatedwork}
\subsection{Google File System}
Google File System (GFS) \cite{gfs} is designed to provide efficient and reliable access to data using clusters of commodity machines located in geographically spread data-centers. In particular, GFS aims to provide high data throughput, low latency, and reliability in the face of individual server failures. Because GFS runs over commodity hardware, the proposed model considers machine and network faults as being the norm rather than the exception \cite{gfs}.  This assumption is typical in modern distributed systems of any kind. GFS is the basis for an open-source project, which has seen continuous improvements over the years, called Hadoop Distributed File System (HDFS) \cite{hadoop}; we will not discuss it here because it holds many similarities to GFS with key differences being the approach to security management. One advantage of GFS over other architectures is that it leverages a distributed file system that tailors to the designers observation that, once written, files on google provided services, are only read, often sequentially, with random writes being practically non-existent, shifting optimization focus from latency to throughput, from random writes to appends when it comes to atomicity and no caching of data blocks \cite{gfs}. In the GFS architecture cluster consists of a single master and multiple chunk-servers. The Master maintains a mapping from files to chunks, the current location of chunks, and is responsible for failure detection over the chunk-servers as well as chunk-migration between them. Clients communicate, for reads and writes, with the Master for metadata operations and directly with the chunk-servers for data-bearing operations. Neither the client nor the chunk-servers cache file data because most of google's client applications stream through huge files. A read operation involves obtaining chunk handles and locations from the Master and then retrieving the chunks from one of the many available chunk-severs. For writes, the client splits the file into 64MB chunks and asks the Master to assign them an identifier, and then the client uploads his chunks to the chunk-servers indicated by the Master.

@@

\newpage\subsection{Ceph File System}
Ceph \cite{ceph} is an open-source, storage system that, like GFS, runs over commodity hardware. Ceph aims to deliver reliable, autonomous, distributed object, block, and file storage that is self-healing, self-managing, and has no single point of failure. The base architectural component of Ceph is the RADOS object store, and all content a client reads and writes to the system are represented as objects, irrespective of the objects' original data type. RADOS is for all purposes a Ceph cluster, and its storage space originates from the disks of multiple machines. Ceph architecture is slightly more complicated than that of GFS, so we will summarize the components that are of interest to us. Inside a cluster, we find three types of machines, storage devices, abstracted with a layer called Object Software Daemon (OSD), Monitors, and Metadata Servers. A small count of Monitors maintain cluster membership and state using PAXOS \cite{paxos} to make consensual decisions. Metadata Servers only required sharing the distributed file system, are used for POSIX semantics, e.g., directory hierarchy and file metadata such as owner and last accessed timestamps. Monitors can detect if metadata servers are down and launch new ones as required. Finally, OSDs store clients' data, and there are anywhere between dozens to thousands of these in a Ceph cluster.

@@

\newpage\section{P2P Overlays}
\subsection{Structured Overlays}
\subsubsection{Kademlia} \cite{kademlia} is structured overlay, yet it is fully decentralized, designed for key-value storage systems, which, according to the authors, combines provable consistency, performance, latency-minimizing routing, and symmetric topology \cite{kademlia}. Kademlia peer-identifiers and data-keys are built as in Chord\cite{chord}, using consistent hashing. Inserts and lookups on the overlay are unidirectional, according to the distance between two identifiers, calculated using XOR. The unidirectionality allows caching to be done along lookup paths, alleviating hot spots. Each peer in the overlay maintains a binary tree, routing table, whose leaves are lists of peers, called k-buckets, at distances $2^{x}$ to $2^{x+1}$, $x \in\hspace{1mm}\mathclose[0,160\mathclose]$ from itself, sorted according to a least recently seen policy. As a peer receives messages if the sender is a new acquaintance, that sender is placed on the proper k-bucket if there is enough space in it; otherwise, it will replace the least seen peer only if that peer fails to respond to a ping. By preferring old acquaintances, Kademlia seeks to enhance the stability of the overlay based on the observation that the longer a peer remains connected to the network, the higher the probability that it will remain connected at least another hour\cite{ssaroiu:msp2pfss}. This policy also prevents some denial of service attacks, where an enemy floods the network with new peers. The XOR metric gives Kademlia an advantage over other DHT-based implementations by making routing tables less rigid and by allowing peers in the network to learn useful routing information from discovery and lookup messages. Because of the existence of k-buckets, routing can also be done around failed nodes, providing some resistance to churn and partitioning.

@@

\subsubsection{Self-Chord} \cite{selfchord} is a bio-inspired, structured overlay, specifically designed to be used in cloud computing or grid computing architectures. Self-chord foundations are that of the standard Chord, both in terms of identifiers and the way peers organize themselves on the overlay, i.e., in ring-like shapes. However, it has a few benefits, which are the result of having mobile agents, called ants, independently roaming the overlay network, but collaboratively re-arranging resource identifiers so that similar data items are clustered around neighboring peers with relatively proportional workloads, meaning that data item placement is not based solely on the numerical-space shared between them and possible placement peers \cite{sotart, selfchord}. Hence advantages of Self-Chord over the standard implementation include lack of need to assign data to well-specified peers since keys are independently defined, leading to better behavior under high churn conditions. Less overlay maintenance overhead since when a peer joins or leaves the network, ants will autonomously and eventually, distribute data items appropriately; Lastly, since load-balancing is continuously being carried out by ants, the scalability and robustness of the system increases since individual peer failures are less likely to lead to massive file loss. According to the authors, the behavior of the ants ensures that even under highly dynamic, unfavorable conditions, rearrangement, and discovery of data items takes only logarithmic time.

@@

\subsection{Unstructured Overlays}
\subsubsection{Gia} \cite{gia} is a decentralized overlay protocol whose purpose is to overcome the drawbacks of Gnutella\cite{gnutella-rfc, gnutella-cs}. Gnutella is the first-ever decentralized, unstructured P2P protocol. Gnutella's main issues are lack of scalability, as resource discovery bases itself on simple message flooding mechanisms and how easily some peers become overloaded when faced with high rates of aggregate queries, leading to system degradation. Gia proposes the use of super-nodes, which are now also part of the Gnutella protocol. The super-nodes receive and route queries to peers holding data, but these super-nodes and the construction of the topology around them is dynamic and adaptive. Gia recognizes the heterogeneity and resource constraints of the peers and also uses random walks to alter the topology adaptively as well as lookup operations. The results show that Gia is three to five orders of magnitude better than that of Gnutella. Random-walks by themselves, while minimizing the network overhead, are less likely to find appropriate responses to the performed lookup, unless they are biased to high-degree peers, but this can make peer overloading more frequent. Thus, to tackle both problems, Gia implements a topology adaptation that puts most peers within reach of high capacity peers, while simultaneously ensuring that they can handle the likely incoming requests. Lastly, all peers keep pointers to data items kept by their immediate one-hop neighbors.

@@

\subsubsection{BlatAnt} \cite{blatant} is a bio-inspired, unstructured overlay, designed for grid computing architectures where resource discovery needs to be efficient; to this end, BlatAnt focuses solely on bounded-diameter optimization in order to minimize network overhead and lookup latency, by using swarm intelligence to support flood-like discovery protocols. Like in Self-Chord\cite{selfchord}, ant agents are used to optimize and maintain a self-structured P2P overlay, conversely, in BlatAnt the ants do not change the location of data items from peer to peer, but instead build and optimize the overlay such that the network uses a minimal number of connections between peers and such that the diameter of the overlay is bounded. The idea is that by doing so, the average number of hops, i.e., average path length, required to obtain a data item from a lookup will be smaller, thus achieving the sought goals. BlatAnt also uses local index caching to improve efficiency further; The authors argue that local caches,  as an alternative to semantic clustering such as the ones used in Antares\cite{antares} and Self-Chord, are better for self-organized overlays where a stable topology can not be guaranteed. The overlay management can be viewed in three major components.

@@

\newpage\section{Membership Management Protocols}
\subsection{CYCLON} CYCLON\cite{cyclon} is a protocol that emerges as an inexpensive solution to the maintenance of gossip-based unstructured overlays. CYCLON is useable even in highly dynamic environments, providing peer degree symmetry, low diameter, low clustering, resilience to churn and massive node failures, in a fully decentralized manner. One fundamental issue of unstructured networks is to avoid partitioning. One naive solution would be for each peer to keep a view of all the peers in the network and create connections has needed, but that solution is not scalable.

@@

We conclude by reinforcing that CYCLON is a scalable and robust solution for peer-sampling, providing good self-healing and self-organization capabilities without complex algorithms or apparent downsides other than the one where too few peers in the system may cause it to be under performant. In a way, it relates to BlatAnt since the proposal passively bounds the diameter of the network.

\subsection{Newscast}
Newscast \cite{newscast} is an alternative peer-sampling protocol

\section{Cloud-Assisted P2P Networks}

Many academic and commercial projects have been developed over the years where cloud, grid, and other centralized architectures are partly assisted by P2P networks. In these systems, peers are usually machines belonging to the clients of the service, and they help to decrease the load on the system, hence reducing its cost. This approach is especially prevalent in streaming and online gaming services, for example, Blizzard Activision uses their player base to help distribute installers and patches in a P2P fashion, and Digital Extremes uses, in their free-to-play game Warframe, P2P networks in particular, but not all, group missions where individuals are in instances of the game inaccessible by players outside the group. However, only recently, was the alternative explored, where P2P networks serve the vast majority of the service and are assisted by stable cloud services only to help guarantee some quality of service property of the system under critical situations. We present two of them.

\subsection{CLOUDCAST}
CLOUDCAST \cite{cloudcast} is a P2P-assisted cloud architecture that differs from other proposals because it does not offload the provided service to peers only when service availability is not affected. Furthermore, the architecture does not augment P2P systems with elastic computing nodes that perform particular tasks beyond the reach of the P2P network, such has bulk-synchronous content distribution. CLOUDCAST functionality depends solely on a passive storage service that transparently integrates into a gossip-based P2P system. The reasoning behind not using elastic computing instances is to eliminate fixed renting costs. Like that, the focus is on minimizing the costs associated with storage and bandwidth. The architecture uses two gossip protocols, one for membership management and another for information dissemination. Key results of the proposal include: accesses to the cloud service are bounded regardless of the number of clients in the system; the cloud service only distributes content when few peers are available; the overlay does not suffer from partitioning regardless of the number of peers in the system; the two illustrated use cases have an absurdly low yearly cost when compared to pure cloud approaches, and finally the authors recognize that the vast part of their bill is tied to membership management and not to actual storage costs, thus using strategies that adapt to the size of the network could yield even better savings.

The passive cloud storage service is a key-value store that can be accessed via GET and PUT operations, but it cannot under any circumstance initiate communications with the system peers. Two problems are identified, \textit{membership management} and \textit{information diffusion}. To solve the first problem, CLOUDCAST uses the CYCLON protocol to do peer-sampling. From other alternatives, CYCLON was chosen not for its particular inexpensiveness, but because the protocol maintains a random overlay topology that works even during high churn periods and even in the presence of catastrophic failures where up to 80\% of the peers fail. CYCLON properties also guarantee that references to the cloud server are not too many keeping usage cost under control. Finally, the cloud server can easily integrate in the protocol, in particular, CYCLON does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; Finally, the cloud server can easily integrate into the protocol, in particular, CYCLON does not require the cloud server to initiate any communications, i.e., its active roles can be carried out by peers interacting with it; In this regard, the dummy process carried by the peers occasionally refreshes pointers to the cloud to avoid losing all references to it in the system. To deal with the case where all references cloud are lost due to the dynamic characteristics of the environment, e.g. churn and lost messages, all peers keep a logical timestamp of the last round they heard about the cloud server updating it whenever they successfully contact it or because one of their a \textit{neighbor} heard about it more recently then them, if after \textit{t} rounds peer has not updated this timestamp he will create a new cloud reference with probability $p\textsubscript{recovery}$.

The second problem is solved using two different implementations of gossip algorithms explained in depth in \cite{epidemic_algorithms}. They are a \textit{rumor-mongering, coin and blind} approach which favors fast news spreading and consists in having a peer \textit{i} periodically forwarding updates he learns about to some other peer \textit{j}, then deciding with probability $p\textsubscript{rumor}$ if he should, permanently, stop gossiping the update; And a \textit{anti-entropy, push-pull} approach which guarantees that news are eventually known by all participants of the system and consists in having each peer \textit{i} periodically contacting a random neighbor \textit{j} to exchange update they know about. All updates to the data are recorded in the cloud server through a PUT operation before becoming hot-topics, thus triggering rumor-mongering. Peers never contact the cloud server for rumor-mongering purposes. The anti-entropy approach can be piggybacked to the CYCLON peer-sampling step, thus decreasing message overhead; the authors, however, do not mention this optimization.

\subsection{INSERT ANOTHER}
\section{Block-Level Data Replication}
\section{Trust Models}
\section{Probabilistic Swarm Guidance}
%%%%%%%%%%%%%%%%%%%%
% END RELATED WORK %
%%%%%%%%%%%%%%%%%%%%
